Assessing spatial correlations in pulsar timing arrays with the noise-marginalized optimal statistic

### 2018-04-29 manuscript/manuscript.tex
added P-P plot

In Fig.~\ref{fig:os_datasetstats} we show the optimal statistic 
for 200 different realizations of a GWB with $\Agw = 5\times10^{-15}$ 
computed using the three techniques described above. 
Using the noise values from individual noise analyses 
systematically underestimates the strength of the GWB, 
while using the noise values from a common noise analysis 
more accurately recovers the injected value. 
The fixed-noise analysis using the individual noise results finds 
$\hat{A}^2 = (7.9 \pm 6.8) \times10^{-30}$ and $\rho = 2.3 \pm 1.5$, 
averaging over realizations of the GWB. 
The fixed-noise and noise-marginalized analyses 
using the common noise results both give 
$\hat{A}^2 = (2.4\pm1.2)\times10^{-29}$ and $\rho = 4.1 \pm 1.7$.
\begin{figure}[t]
	\includegraphics[width=0.95\columnwidth]{plots/optstat_A5e-15.pdf}
	\caption{Optimal statistic and SNR for 200 simulated datasets 
			containing an injected GWB. 
			The dashed vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
			We show the values found by fixing the red noise 
			based on the results from individual and common noise analyses 
			(blue and orange lines, respectively) 
			as well as the mean optimal statistic and mean SNR found by marginalizing over 
			the red noise parameters using the results of a common noise analysis (green lines). 
			Using the red noise values from individual noise analyses leads to values of $\hat{A}^2$ 
			that are significantly lower than the injected value, 
			while using the values from a common noise analysis gives values of $\hat{A}^2$ 
			that are closer to the injected value.}
	\label{fig:os_datasetstats}
\end{figure}

The fixed-noise and noise-marginalized analyses give the same results 
for $\Agw = 5\times10^{-15}$, 
but there are differences between them when analyzing datasets 
containing smaller injected values of $\Agw$. 
%Figure~\ref{fig:os_datasetstats2} shows the mean optimal statistic and mean SNR 
%for 100 realizations of a GWB with $\Agw = 10^{-15}$. 
%The fixed-noise analysis using the individual noise values 
%gives a mean value of $\hat{A}^2 = 2.9\times10^{-31}$ and a mean SNR of 1.1. 
%The fixed-noise analysis using the common noise values 
%gives a mean value of $\hat{A}^2 = 8.4\times10^{-31}$ and a mean SNR of 2.1, 
%which the noise-marginalized analysis gives a mean value of 
%$\hat{A}^2 = 1.1\times10^{-30}$ and a mean SNR of 2.4.
%\begin{figure}[t]
%	\includegraphics[width=0.95\columnwidth]{plots/optstat_A1e-15.pdf}
%	\caption{Optimal statistic and SNR for 100 simulated datasets 
%			containing an injected GWB with $\hat{A}^2 = 10^{-30}$ (dashed vertical line). 
%			We show the values found by fixing the red noise 
%			based on the results from individual and common noise analyses 
%			(blue and orange lines, respectively) 
%			to the mean optimal statistic and mean SNR found by marginalizing over 
%			the red noise parameters using the results of a common noise analysis (green lines). 
%			Using the red noise values from individual noise analyses leads to values of $\hat{A}^2$ 
%			that are significantly lower than the injected value, 
%			while using the values from a common noise analysis gives values of $\hat{A}^2$ 
%			that are closer to the injected value.}
%	\label{fig:os_datasetstats2}
%\end{figure}
We also compare these three methods 
based on how far the computed values of the optimal statistic are from the injected value of of the GWB 
compared to the uncertainty in the optimal statistic,
\begin{equation}
	s = \frac{\hat{A}^2 - \Agw^2}{\sigma} = \frac{\hat{A}^2 - \Agw^2}{\hat{A}^2/\rho} \,.
\end{equation}
Figure~\ref{fig:os_compare} shows the distribution of $(\hat{A}^2-\Agw^2)/\sigma$
\begin{figure*}[t]
	\includegraphics[width=0.95\columnwidth]{plots/pp_plot.pdf}
	\caption{\sv{write caption}}
	\label{fig:os_compare}
\end{figure*}

>>>>

In Fig.~\ref{fig:os_datasetstats} we show the optimal statistic 
for 200 different realizations of a GWB with $\Agw = 5\times10^{-15}$ 
computed using the three techniques described above. 
Using the noise values from individual noise analyses 
systematically underestimates the strength of the GWB, 
while using the noise values from a common noise analysis 
more accurately recovers the injected value. 
The fixed-noise analysis using the individual noise results finds 
$\hat{A}^2 = (7.9 \pm 6.8) \times10^{-30}$ and $\rho = 2.3 \pm 1.5$, 
averaging over realizations of the GWB. 
The fixed-noise and noise-marginalized analyses 
using the common noise results both give 
$\hat{A}^2 = (2.4\pm1.2)\times10^{-29}$ and $\rho = 4.1 \pm 1.7$.
\begin{figure}[htb]
	\includegraphics[width=0.95\columnwidth]{plots/optstat_A5e-15.pdf}
	\caption{Optimal statistic and SNR for 200 simulated datasets 
			containing an injected GWB. 
			The dashed vertical line indicates the injected value, $\hat{A}^2 = 2.5 \times 10^{-29}$. 
			We show the values found by fixing the red noise 
			based on the results from individual and common noise analyses 
			(blue and orange lines, respectively) 
			as well as the mean optimal statistic and mean SNR found by marginalizing over 
			the red noise parameters using the results of a common noise analysis (green lines). 
			Using the red noise values from individual noise analyses leads to values of $\hat{A}^2$ 
			that are significantly lower than the injected value, 
			while using the values from a common noise analysis gives values of $\hat{A}^2$ 
			that are closer to the injected value.}
	\label{fig:os_datasetstats}
\end{figure}

The fixed-noise and noise-marginalized analyses give the same results 
for $\Agw = 5\times10^{-15}$, 
but there are differences between them when analyzing datasets 
containing smaller injected values of $\Agw$. 
%Figure~\ref{fig:os_datasetstats2} shows the mean optimal statistic and mean SNR 
%for 100 realizations of a GWB with $\Agw = 10^{-15}$. 
%The fixed-noise analysis using the individual noise values 
%gives a mean value of $\hat{A}^2 = 2.9\times10^{-31}$ and a mean SNR of 1.1. 
%The fixed-noise analysis using the common noise values 
%gives a mean value of $\hat{A}^2 = 8.4\times10^{-31}$ and a mean SNR of 2.1, 
%which the noise-marginalized analysis gives a mean value of 
%$\hat{A}^2 = 1.1\times10^{-30}$ and a mean SNR of 2.4.
%\begin{figure}[t]
%	\includegraphics[width=0.95\columnwidth]{plots/optstat_A1e-15.pdf}
%	\caption{Optimal statistic and SNR for 100 simulated datasets 
%			containing an injected GWB with $\hat{A}^2 = 10^{-30}$ (dashed vertical line). 
%			We show the values found by fixing the red noise 
%			based on the results from individual and common noise analyses 
%			(blue and orange lines, respectively) 
%			to the mean optimal statistic and mean SNR found by marginalizing over 
%			the red noise parameters using the results of a common noise analysis (green lines). 
%			Using the red noise values from individual noise analyses leads to values of $\hat{A}^2$ 
%			that are significantly lower than the injected value, 
%			while using the values from a common noise analysis gives values of $\hat{A}^2$ 
%			that are closer to the injected value.}
%	\label{fig:os_datasetstats2}
%\end{figure}
We also compare these three methods

Figure~\ref{fig:os_compare} shows the distribution of $(\hat{A}^2-\Agw^2)/\sigma$
\begin{figure*}[t]
	\includegraphics[width=0.95\textwidth]{plots/pp_plot.pdf}
	\caption{P--P plot showing the cumulative fraction of simulations for which $\Agw^2$ lies within 
			a given confidence interval of the measured $\hat{A}^2$. 
			The probability distribution of $\hat{A}^2$ is assumed to be a Gaussian 
			with variance $\sigma^2_{\hat{A}^2}$. 
			The fixed-noise optimal statistic using the individual and common noise results 
			both give biased values of $\hat{A}^2$, particularly for small values of \Agw, 
			while the noise-marginalized optimal statistic gives more accurate values of $\hat{A}^2$ 
			over a large range of injected values of \Agw.}
	\label{fig:os_compare}
\end{figure*}
